---
editor_options:
  chunk_output_type: console
output:
  word_document: default
  html_document: default
---
```{r}
library(tidyverse)
library(cluster)
library(factoextra)
library(dendextend)
```

```{r}
trucks <- read_csv("trucks.csv")
```

Task 1
```{r}
ggplot(data = trucks, aes(Distance, Speeding)) +
  geom_point()
```
Drivers who average 0-100 miles a day tend to speed 0-65% of their time driving. Drivers who average 125-250 miles a day tend to speed more often 0-25% of the time, and less often 25-100% of the time. When someone is driving many miles in one day, they are more likely to speed for more of their trip.

Task 2
```{r}
trucks2 <- trucks %>%
  select(Distance, Speeding)
  as.data.frame(scale(trucks2))
```

Task 3
```{r}
set.seed(1234)
clusters <- kmeans(trucks2, 2)
fviz_cluster(clusters, trucks2)
```
ANSWER

Task 4
```{r}
set.seed(123)
fviz_nbclust(trucks2, kmeans, method = "wss")

set.seed(123)
fviz_nbclust(trucks2, kmeans, method = "silhouette")
```
WSS: 2
Silhouette: 2

Task 5
```{r}
set.seed(1234)
clusters1 <- kmeans(trucks2, 2)
clusters1 
fviz_cluster(clusters1, trucks2)
```

Task 6
It doesn't look like these clusters share many characteristics. There are two well defined groups.

```{r}
wine <- read_csv("wineprice.csv")

wine2 <- wine %>%
  select(Price, WinterRain, AGST, HarvestRain, Age)
  as.data.frame(scale(wine2))
```

Task 7
```{r}
set.seed(123)
clusters <- kmeans(wine2, 2)
fviz_cluster(clusters, trucks2)

set.seed(123)
fviz_nbclust(wine2, kmeans, method = "wss")

set.seed(123)
fviz_nbclust(wine2, kmeans, method = "silhouette")
```
WSS: 2
Silhouette: 2

Task 8
```{r}
set.seed(1234)
clusters2 <- kmeans(wine2, 2)
clusters2 
fviz_cluster(clusters2, wine2)
```

Task 9
```{r}
m = c( "average", "single", "complete", "ward")
names(m) = c( "average", "single", "complete", "ward")

ac = function(x) {
  agnes(wine2, method = x)$ac
}
map_dbl(m, ac)
```

```{r}
hc = agnes(wine2, method = "ward")
pltree(hc, cex = 0.6, hang = -1, main = "Agglomerative Dendrogram") 
```

Task 10
```{r}
hc2 = diana(wine2)
pltree(hc2, cex = 0.6, hang = -1, main = "Divisive Dendogram")
```

```{r}
plot(hc, cex.axis= 0.5) 
rect.hclust(hc2, k = 5, border = 2:6)
```

```{r}
plot(hc2, cex.axis= 0.5) 
rect.hclust(hc2, k = 5, border = 2:6)
```

