---
output:
  word_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---
```{r}
#install.packages("nnet")
#install.packages("caretEnsemble")
```

```{r}
library(tidyverse)
library(caret)
library(nnet)
library(rpart)
library(caretEnsemble)
library(ranger)
```

```{r}
parole <- read_csv("parole.csv")
```

```{r}
parole <- parole %>%
  mutate(male = as_factor(as.character(male))) %>%
  mutate(male = fct_recode(male,
    "Female" = "0",
    "Male" = "1"))
```

```{r}
parole <- parole %>%
  mutate(race = as_factor(as.character(race))) %>%
  mutate(race = fct_recode(race,
    "white" = "1",
    "poc" = "2")) #person of color
```

```{r}
parole <- parole %>%
  mutate(state = as_factor(state)) %>%
  mutate(state = fct_recode(state,
    "Other" = "1",
    "Kentucky" = "2",
    "Louisiana" = "3",
    "Virginia" = "4"))
```

```{r}
parole <- parole %>%
  mutate(crime = as_factor(as.character(crime))) %>%
  mutate(crime = fct_recode(crime,
    "other" = "1",
    "larceny" = "2",
    "drug-related" = "3",
    "driving-related" = "4"))
```

```{r}
parole <- parole %>%
  mutate(multiple.offenses = as_factor(as.character(multiple.offenses))) %>%
  mutate(multiple.offenses = fct_recode(multiple.offenses,
    "single_offense" = "0",
    "multiple_offenses" = "1"))
```

```{r}
parole <- parole %>%
  mutate(violator = as_factor(as.character(violator))) %>%
  mutate(violator = fct_recode(violator,
    "no_violations" = "0",
    "violated" = "1"))
```

Task 1
```{r}
set.seed(12345) 
train.rows <- createDataPartition(y = parole$violator, p=0.7, list = FALSE)
train <- parole[train.rows,] 
test <- parole[-train.rows,]
```

Task 2
```{r}
fitControl <- trainControl(method = "cv", number = 10)
nnetGrid <- expand.grid(size = 12, decay = 0.1)
set.seed(1234)
nnet <- train(x=as.data.frame(parole[,-1]),y=as.matrix(parole$violator), 
                 method = "nnet",
                 trControl = fitControl,
                 tuneGrid = nnetGrid,
                 verbose = FALSE)

```

Task 3
```{r}
predNet <- predict(nnet, train)
confusionMatrix(predNet, train$violator, positive = "violated")
```
Accuracy: 1
Sensitivity: 1
Specificity: 1
No information rate: .8837

Task 4
```{r}
fitControl <- trainControl(method = "cv", number = 10)
nnetGrid1 <- expand.grid(size = seq(from = 1, to = 12, by = 1), decay = seq(from = .01, to = 0.5, by = 0.1))
set.seed(1234)
nnet1 <- train(x=as.data.frame(parole[,-1]),y=as.matrix(parole$violator), 
                 method = "nnet",
                 trControl = fitControl,
                 tuneGrid = nnetGrid1,
                 verbose = FALSE,
                 trace = FALSE)

```

Task 5
```{r}
predNet1 <- predict(nnet1, train)
confusionMatrix(predNet1, train$violator, positive = "violated")
```
Accuracy: 1
Sensitivity: 1
Specificity: 1
No information rate: .8837

Task 6
```{r}
predNet2 <- predict(nnet, test)
confusionMatrix(predNet2, test$violator, positive = "violated")
```
Accuracy: 1
Sensitivity: 1
Specificity: 1
No information rate: .8861

Task 7
```{r}
predNet3 <- predict(nnet1, test)
confusionMatrix(predNet3, test$violator, positive = "violated")
```
Accuracy: 1
Sensitivity: 1
Specificity: 1
No information rate: .8861

Task 8
There is no difference in accuracy between the two models, but the confusion matrix shows that the largest class forms the majority of the population.

Task 9
```{r}
ensemblecontrol = trainControl(
  method = "cv",
  number = 5,
  savePredictions = "final",
  classProbs = TRUE,
  summaryFunction = twoClassSummary
  )
```

```{r}
set.seed(111)
model_list = caretList(
  x = as.data.frame(train[,-9]),y=as.matrix(train$violator),
  metric = "ROC",
  trControl= ensemblecontrol,
  methodList=c("glm","ranger","rpart", "nnet"),
  tuneList=list(
    rf = caretModelSpec(method = "ranger", tuneLength=6),
    rpart = caretModelSpec(method = "rpart", tuneLength=6),
    nn = caretModelSpec(method = "nnet", tuneLength=6, trace=FALSE)
  ))
```

```{r}
as.data.frame(predict(model_list, newdata=head(train)))
```

```{r}
modelCor(resamples(model_list))
```
Rf and ranger are the highest correlated models.

```{r}
ensemble = caretEnsemble(
  model_list, 
  metric="ROC",
  trControl=trainControl(
    method = "cv",
    number= 5,
    summaryFunction=twoClassSummary,
    classProbs=TRUE
    ))
```

```{r}
summary(ensemble)
```
From the summary, we see that the resulting ROC for the ensemble is .8106. This is a higher value than the ROC for the rpart model and a lower value than the ROC for the rf, rpart, nn, glm, and ranger models.

```{r}
pred_ensemble = predict(ensemble, train, type = "raw")
confusionMatrix(pred_ensemble,train$violator)

pred_ensemble_test = predict(ensemble, test, type = "raw")
confusionMatrix(pred_ensemble_test,test$violator)
```
Train
Accuracy: .9619
Sensitivity: 1
Specificity: .6727

Test
Accuracy: .8911
Sensitivity: .9888
Specificity: .1304

Task 10

```{r}
stack = caretStack(
  model_list,
  method ="glm",
  metric ="ROC",
  trControl = trainControl(
    method = "cv", 
    number = 5,
    savePredictions = "final",
    classProbs = TRUE,
    summaryFunction = twoClassSummary
  ))

```

```{r}
print(stack)
```
ROC .811

```{r}
pred_stack = predict(stack, train, type = "raw")
confusionMatrix(pred_stack,train$violator)

pred_stack_test = predict(stack, test, type = "raw")
confusionMatrix(pred_stack_test,test$violator)
```
Train
Accuracy: .9619
Sensitivity: 1
Specificty: .6727

Test
Accuracy: .8911
Sensitivity: .988
Specificity: .1304
